---
title: "Interference"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(statnet)
library(ergm)
library(stringr)
library(dplyr)
library(purrr)
library(tidyr)
library(data.table)
library(boot)
library(MatchIt)
library(Matching)
library(igraph)
library(ggplot2)
```

```{r}
source('NeighGPS2stage.subclass.opts.Bootstrap.R')
source('NeighGPSfunctions.subclass.opts.Bootstrap.R')
```



```{r}
dat=read.csv('facebook_combined.txt',header=F,sep=' ')
g=graph.data.frame(dat,directed=FALSE)
#plot(g,vertex.cex=0.1)
```

```{r}
vcount(g)
```

```{r}
plot(degree_distribution(g,cumulative = F))
```


```{r}
network.features = list(); j =1;
for(i in c(0,107,348,414,686,698,1684,1912,3437,3980)){
  ego0 = read.csv(paste0('facebook/',i,'.egofeat'),sep=' ',header = F)
  feature0 = read.csv(paste0('facebook/',i,'.feat'),sep=' ',header = F)
  feature0.name = read.csv(paste0('facebook/',i,'.featnames'),sep=' ',header = F)
  network1.features = rbind(ego0,feature0)
  colnames(network1.features)[2:dim(network1.features)[2]]<- feature0.name %>% 
    mutate(splits = str_split(V2,';')) %>% 
    rowwise() %>% 
    mutate(
      identifier = paste(c(splits[-length(splits)],V4),collapse='_')
    ) %>% .$identifier
  network.features[[j]] <- network1.features
  j <- j + 1;
}

```


```{r}
rbind.all.columns <- function(x, y) {
 
    x.diff <- setdiff(colnames(x), colnames(y))
    y.diff <- setdiff(colnames(y), colnames(x))
 
    x[, c(as.character(y.diff))] <- 0
 
    y[, c(as.character(x.diff))] <- 0
 
    return(rbind(x, y))
}
networks = network.features %>% reduce(rbind.all.columns) %>% filter(V1 %in% get.vertex.attribute(g,'name')) %>% filter (! duplicated(V1))
networks = networks[order(networks$V1),]
```


```{r}
dat$V1 = dat$V1 + 1
dat$V2 = dat$V2 + 1
fbnet <- as.network(x = dat, # the network object
                  directed = F, # specify whether the network is directed
                  loops = FALSE, # do we allow self ties (should not allow them)
                  matrix.type = "edgelist", # the type of input
                  vertex.attrnames = 1:4039
)
networks$V1 = 1:4039


education_type = networks %>% 
  dplyr::select(V1,education_type_53:education_type_55) %>%
  mutate(education_type = paste0(education_type_53,education_type_54,education_type_55)) %>%
  .$education_type
education_type = as.numeric(education_type) %>% recode('10'=2,'11'=3,'100'=4,'101'=5,'110'=6,'111'=7,'0'=0,'1'=1) 
network::set.vertex.attribute(fbnet,'Education_type',education_type)

gender = networks %>% 
  dplyr::select(V1,starts_with('gender')) %>%
  mutate(gender = paste0(gender_77,gender_78)) %>%
  .$gender
gender = as.numeric(gender) %>% recode('10'=2,'1'=1)
network::set.vertex.attribute(fbnet,'gender',gender)


ht = networks %>% 
  dplyr::select(starts_with('birthday')) 
  
age = lapply(colnames(ht)[apply(ht,1,which.max)],function(x) str_split(x, '_',simplify = T)[length(str_split(x, '_',simplify = T))]) %>% unlist()
age = as.numeric(age) %>% recode("0"=0,'1'=1,'2'=2,'208'=3,'209'=4,'210'=5,'3'=6,'377'=7,'4'=8,'5'=9,'6'=10,'7'=11,.default =12)
network::set.vertex.attribute(fbnet,'age',age)

ht = networks %>% 
  dplyr::select(starts_with('hometown_id')) 

hometown = lapply(colnames(ht)[apply(ht,1,which.max)],function(x) str_split(x, '_',simplify = T)[length(str_split(x, '_',simplify = T))]) %>% unlist()
hometown = as.numeric(hometown) %>% recode("79"=0,'84'=1,'908'=2,'935'=3,'84'=4,.default =5)
network::set.vertex.attribute(fbnet,'hometown',hometown)

```


```{r}
par(mfrow=c(1,1)) # Back to 1-panel plots
plot(fbnet, vertex.col='gender')
legend('bottomleft',fill=1:2,legend=paste('gender',1:2),cex=0.75)
```


```{r}
par(mfrow=c(1,1)) # Back to 1-panel plots
plot(fbnet, vertex.col='age')
legend('bottomleft',fill=0:12,legend=paste('age',0:12),cex=0.75)
```


## Scenario 2 



```{r}
g = graph.data.frame(dat,directed=FALSE,vertices=1:4039)
G = NULL;Friend.Gender = NULL;Friend.Hometown = NULL;Friend.Education = NULL;Friend.Age = NULL;dg = NULL
N=4039
for(i in V(g)){
  neigh = neighbors(g,v = i)
  Friend.Gender[i] = mean(gender[neigh])
  Friend.Age[i] = mean(age[neigh])
  dg[i] = mean(degree(g,v=i))
}
#beta1=-4.5
#beta2=0.4
#beta3=0.1
#beta4=2.5
#beta5=0.3

beta1=-5
beta2=0.5
beta3=0.1
beta4=2.5
beta5=0.4
#
#beta1=-60
#beta2=5
#beta3=1
#beta4=25
#beta5=4

set.seed(123)
Z_prob = exp(beta1+beta2*gender+beta3*age+beta4*Friend.Gender+beta5*Friend.Age)/(1+exp(beta1+beta2*gender+beta3*age+beta4*Friend.Gender+beta5*Friend.Age))
Z = map_dbl(1:length(gender),function(x){rbernoulli(n = 1,Z_prob[x])})

ggplot(data.frame(pps = Z_prob, trt = as.integer(Z)), aes(x=pps,group=trt,color=trt)) +
  geom_density() +
  ggtitle('propensity scores')
#ggsave(filename = paste0(beta1,"_",beta2,"_",beta3,"_",beta4,"_",beta5,".pdf"))

for (i in 1:N){
  neigh = neighbors(g,v = i)
    G[i] = mean(Z[neigh])
}
     
c(mean(G[Z==T],na.rm = T),mean(G[Z==F],na.rm = T),mean(G,na.rm = T))

```





  
  
```{r}
Zcovariates = data.frame(Z,gender,age,Friend.Gender,Friend.Age,G,dg)
Zcovariates = Zcovariates[!is.na(Zcovariates$G),]
covariateBalance = Zcovariates %>% group_by(Z) %>% summarise_all(mean) %>% t()
d = dim(covariateBalance)[1]
covariatestd = Zcovariates  %>% group_by(Z) %>% summarise_all(var) %>% t()
covariate = data.frame(cbind(covariateBalance[2:d,c(2,1)],(covariateBalance[2:d,2] - covariateBalance[2:d,1])/sqrt(covariatestd[2:d,1]+covariatestd[2:d,2])))
colnames(covariate) <- c('Treated','Control','Stand.Diff')
#rownames(covariate) <- c('Gender','Hometown','Friends.Hometown','Friends.Gender','G')
knitr::kable(covariate,caption = 'Covariate Balance across individual treatment arms')
```

```{r}
covariateBalance = Zcovariates %>% 
  mutate(G = ifelse(G>=0.5,1,0)) %>%
  group_by(G) %>% summarise_all(mean) %>% t()
covariatestd = Zcovariates  %>% group_by(Z) %>% summarise_all(var) %>% t()
covariate = data.frame(cbind(covariateBalance[2:d,c(2,1)],(covariateBalance[2:d,2] - covariateBalance[2:d,1])/sqrt(covariatestd[2:d,1]+covariatestd[2:d,2])))
colnames(covariate) <- c('Treated','Control','Stand.Diff')
#rownames(covariate) <- c('Z','Gender','Hometown','Friends.Hometown','Friends.Gender')
knitr::kable(covariate,caption = 'Covariate balance across dichotomized neighborhood treatment arms')
```


```{r}
set.seed(123)
Zcovariates = data.frame(Z,gender,age,Friend.Gender,Friend.Age,G,dg)
covariateBalance = Zcovariates %>% 
  mutate(G = rbernoulli(1,G), Z= as.numeric(Z)) %>% tbl_df()

model1 = glm(G~.-Friend.Age-Friend.Gender,family = binomial(link='logit'),data = covariateBalance)
summary(model1)
model1 = glm(G~.,family = binomial(link='logit'),data = covariateBalance)
summary(model1)
psi=predict(model1,newdata = covariateBalance,type = "response")

ndata=covariateBalance
ndata$gender=1
ndata$Z=0
abc=NULL

  for (j in 0:7){
    ndata$age=j
    abc=c(abc,predict(model1,newdata = ndata,type = "response"))
}
abc2=abc
abc1=abc
dfabc=data_frame(g=rep(Zcovariates$G,32),psi=c(abc1,abc2),gender=rep(c("gender1","gender2"),each=N*16),group=rep(rep(1:16,each=N),2))
ggplot(dfabc%>%filter(group==2),mapping = aes(x=g,y=psi))+geom_line()+facet_grid(.~gender)
abcabc=data_frame(g=Zcovariates$G,psi=abc)
ggplot(abcabc,mapping = aes(x=g,y=psi))+geom_line()
length(abc)
length(abc)/N
#colnames(covariate) <- c('Treated','Control','Stand.Diff')
#rownames(covariate) <- c('Z','Race','Grade','Friends.Grade','Friends.Race','Degree')
#knitr::kable(covariate,caption = 'Covariate balance across dichotomized neighborhood treatment arms')
i=1
j=1
length(abc)
```




# Scenario 1   

The individual treatment $Z_i$ is generated depending on individual gender and age

# Scenario 2

## Simulate Outcome

```{r}
N = length(G[!is.na(G)])
Y.potential = array(0,dim=c(N,3))
delta = c(-5,-8,-10)
for(d in 1:length(delta)){
  mu = 15 - 7*(Z_prob[!is.na(G)] >=0.5) - 15*Z[!is.na(G)]+ 3*Z[!is.na(G)]*(Z_prob[!is.na(G)] >=0.5)+delta[d]*G[!is.na(G)]
  for(i in 1:N){
    Y.potential[i,d] = rnorm(1,mu[i],1)
  }
}
# true main effect and spillover effect
c(mean(- 15+ 3*(Z_prob[!is.na(G)] >=0.5)),delta[1]*mean(G[!is.na(G)]))
trueme=mean(- 15+ 3*(Z_prob[!is.na(G)] >=0.5))
```


```{r}
N = length(G[!is.na(G)])
Y.potential = array(0,dim=c(N,3))
delta = c(-5,-8,-10)
for(d in 1:length(delta)){
  mu = 15 - Friend.Age+7*Friend.Gender-10*(Z_prob[!is.na(G)] >=0.7)-10*Z-10*psi+5*G*(Z_prob[!is.na(G)] >=0.7) + 3*Z[!is.na(G)]*(G[!is.na(G)] >=0.7)+delta[d]*G[!is.na(G)]
  for(i in 1:N){
    Y.potential[i,d] = rnorm(1,mu[i],1)
  }
}
# true main effect and spillover effect
mean(- 10+ 3*G)
trueme=mean(- 10+ 3*G)

```




```{r}
Xz.formula<- "gender+age"
Xg.formula<-"gender+age+Friend.Gender+Friend.Age+dg"

formulaZ <- as.formula(paste('Z~', Xz.formula))
formulaG <- as.formula(paste('G~Z+', Xg.formula))

zsim<-c(0,1)
gsim<-(c(0,quantile(Zcovariates$G, probs = seq(0.1, 0.95, length.out=30-1), na.rm=T)))
Nzsim<-length(zsim)
Ngsim<-length(gsim)
grid_val<-list(zsim=zsim, gsim=gsim)
grid_points<-data.frame(z=rep(zsim,Nzsim),g=rep(gsim,rep(Nzsim,Ngsim))) 

Zcovariates$Y<-Y.potential[,1]

ylim<-c(-11,17)
```


Dose Function 

```{r}
Zcovariates$GG<-Zcovariates$G
data = Zcovariates%>%dplyr::select(-G)
data$Z = as.numeric(data$Z)
formulaY<-as.formula(paste('Y~Z+GG+', Xz.formula))
Ymod<-lm(formulaY, data=data)
mYzg.pois<-matrix(NA, nrow=Nzsim, ncol=Ngsim)
for(i in 1:Nzsim){
	for(j in 1:Ngsim){
  	z<-zsim[i]
	  g<-gsim[j]
	  mYzg.pois[i,j]<-	mean(predict(Ymod, newdata=transform(data, Z=z,GG=g), type='response'))
	}
}
par(mfrow=c(1,2))
plot(gsim, mYzg.pois[1,],type='l', xlab='g', ylab='Y(0,g)', ylim=ylim)
plot(gsim, mYzg.pois[2,],type='l', xlab='g',ylab='Y(1,g)', ylim=ylim)
par(mfrow=c(1,1))

```



```{r}
formulaZ<-as.formula(paste('Z~', Xz.formula))
formulaG<-as.formula(paste('GG~Z+', Xg.formula))
formulaY<-as.formula('Y~Z+GG+I(GG^2)+I(GG^3)+gpsG+I(gpsG^2)+I(gpsG^3)+I(GG*gpsG)')

out.neigh.gps.boot  <- neigh.gps.boot  (Y = Y,
																	 ind.treat = Z,
																	 neigh.treat = GG,
																	 ind.treat_formula = eval(formulaZ),
																	 neigh.treat_formula = eval(formulaG),
																	 outcome_formula = eval(formulaY),
																	 data = data,
																	 grid_val = grid_val,
																	 ind.treat_mod = 'Binomial',
																	 neigh.treat_mod = 'Normal',
																	 outcome_mod = 'lm', #"lm"
                                   outcome_family='Normal',
																	 ind.gps_method='subclass', 
																	 nclasses=5,
                                   pred.full.data=T,
																	 link_function = "logit")


mYzg.gps.boot<-apply(out.neigh.gps.boot$Yzg.bar,c(1,2),mean)
q0.025Yzg<-apply(out.neigh.gps.boot$Yzg.bar,c(1,2),quantile, probs=0.025)
q0.975Yzg<-apply(out.neigh.gps.boot$Yzg.bar,c(1,2),quantile, probs=0.975)

par(mfrow=c(1,2))
plot(gsim, mYzg.gps.boot[1,],type='l', xlab='g', ylab='Y(0,g)', ylim=ylim)
polygon(c(gsim,rev(gsim)),c(q0.025Yzg[1,],rev(q0.975Yzg[1,])),border=NA,col=blues9[3])
lines(gsim,mYzg.gps.boot[1,], type='l')

plot(gsim, mYzg.gps.boot[2,],type='l', xlab='g', ylab='Y(1,g)', ylim=ylim)
polygon(c(gsim,rev(gsim)),c(q0.025Yzg[2,],rev(q0.975Yzg[2,])),border=NA,col=blues9[3])
lines(gsim,mYzg.gps.boot[2,], type='l')
par(mfrow=c(1,1))


# Average Treatment Effect \tau=\int E[Y(1,g)-Y(0,g)]P(G=g)
mdYz<-apply(out.neigh.gps.boot$dYz,1,mean)
q0.025dYz<-apply(out.neigh.gps.boot$dYz,1,quantile, probs=0.025)
q0.975dYz<-apply(out.neigh.gps.boot$dYz,1,quantile, probs=0.975)


# Average Spillover Effect \Delta(z)=\int E[Y(z,g)-Y(z,0)]P(G=g)
mYzg_Yz0<-apply(out.neigh.gps.boot$mYzg_Yz0,1, mean)
q0.025.mYzg_Yz0<-apply(out.neigh.gps.boot$mYzg_Yz0,1,quantile, probs=0.025)
q0.975.mYzg_Yz0<-apply(out.neigh.gps.boot$mYzg_Yz0,1,quantile, probs=0.975)

#Check Balance
#summary(out.neigh.gps.boot$match.mod[[1]])

c(mdYz,mYzg_Yz0)
-13.508447  -5.815126  -5.818051
```



```{r}
formulaZ<-as.formula(paste('Z~', Xz.formula))
formulaG<-as.formula(paste('GG~Z+', Xg.formula))
formulaY<-as.formula('Y~Z+GG+I(GG^2)+I(GG^3)+gpsZ+I(gpsZ^2)+I(gpsZ^3)+gpsG+I(gpsG^2)+I(gpsG^3)+I(GG*gpsG)')
#formulaY<-as.formula('Y~Z+GG+gpsG')
#data = data %>% mutate(GG = G) 
out.neigh.gps.boot  <- neigh.gps.boot  (Y = Y,
																	 ind.treat = Z,
																	 neigh.treat = GG,
																	 ind.treat_formula = eval(formulaZ),
																	 neigh.treat_formula = eval(formulaG),
																	 outcome_formula = eval(formulaY),
																	 data = data,
																	 grid_val = grid_val,
																	 ind.treat_mod = 'Binomial',
																	 neigh.treat_mod = 'Normal',
																	 outcome_mod = 'lm',
                                   outcome_family='Normal',
                                   ind.gps_method='none',
																	 bootstrap = NULL,
                                   pred.full.data=T)


mYzg.gps.boot<-apply(out.neigh.gps.boot$Yzg.bar,c(1,2),mean)

par(mfrow=c(1,2))
plot(gsim, mYzg.gps.boot[1,],type='l', xlab='g', ylab='Y(0,g)', ylim=ylim)
plot(gsim, mYzg.gps.boot[2,],type='l', xlab='g',ylab='Y(1,g)', ylim=ylim)
par(mfrow=c(1,1))

# Average Treatment Effect \tau=\int E[Y(1,g)-Y(0,g)]P(G=g)
dYz<-apply(out.neigh.gps.boot$dYz,1, mean)

# Average Spillover Effect \Delta(z)=\int E[Y(z,g)-Y(z,0)]P(G=g)
mYzg_Yz0<-apply(out.neigh.gps.boot$mYzg_Yz0,1, mean)

c(dYz,mYzg_Yz0)
```




# Main Effect

## Comparison with other approaches

1. A simple difference in mean estimator comparing treated and untreated units;

```{r}
mean(Zcovariates[Zcovariates$Z==1,'Y']) - mean(Zcovariates[Zcovariates$Z==0,'Y'])
```


2. An OLS estimator that regresses the outcome on individual treatment $Z_i$, adjusting only for individual covariates $X_i^{ind}$

```{r}
model = lm(Y~Z + gender + age, data = Zcovariates)
data = model.matrix(model)
#data=Zcovariates
data[,'Z'] = 1
Y.treat = mean(stats::predict(object=model,newdata =data.frame(data[,2:4]),type='response' ))
data[,'Z'] = 0
Y.control = mean(stats::predict(model,newdata =data.frame(data[,2:4]),type='response' ))
Y.treat - Y.control
```

3. An estimator based on a subclassification on the individual propensity score which is estimated using only individual covariates.

```{r}
data_Z = Zcovariates
model = lm(Z~ gender + age, data = data_Z)
data_Z$pps = model$fitted.values
data_Z <- data_Z %>% mutate(block = ntile(pps,5))
data_Z %>%
  group_by(block) %>% 
  summarise_at(funs(mean(.[Z == 1]) -mean(.[Z == 0])),.vars = c('Y')) %>% 
  summarise(m = mean(Y)) %>%
  .$m
```

4. OLS Estimator that regresses the outcome on individual treatment  $Z_i$ adjusting for individual covariates and neighborhood covariates.

```{r}
model = lm(Y~Z + gender + age+Friend.Age + Friend.Gender+dg, data = Zcovariates)
data = model.matrix(model)
data[,'Z'] = 1
Y.treat = mean(stats::predict(model,newdata =data.frame(data),type='response' ))
data[,'Z'] = 0
Y.control = mean(stats::predict(model,newdata =data.frame(data),type='response' ))
Y.treat - Y.control
```

5. An estimator based on a subclassification on the individual propensity score which is estimated using only individual covariates and neighborhood covariates.


```{r}
data_Z = Zcovariates
model = glm(Z~gender + age+Friend.Age + Friend.Gender+dg,family = binomial(link='logit'),data = data_Z)
data_Z$pps=predict(model,newdata = data_Z,type = "response")
data_Z <- data_Z %>% mutate(block = ntile(pps,5))
data_Z %>%
  group_by(block) %>% 
  summarise_at(funs(mean(.[Z == 1])-mean(.[Z == 0])),.vars = c('Y')) %>% 
  summarise(m = mean(Y)) %>%
  .$m

```

```{r}
#trueme=-14.163654
#trueme=-14.163654
results=read.table(file = "res5.txt",stringsAsFactors = FALSE)
res_=apply(results,2,function(x){c(mean(x),sqrt(mean((x-trueme)^2)))})
res_=res_[,-1]
res_df=data.frame(val=c((res_[1,]-trueme),res_[2,]),par=rep(c("Bias","RMSE"),each=24),interference=rep(rep(c("low","medium","high"),each=8),2),estimator=rep(c("Index","Unadjusted","RegressionXind","SubclassXind","RegressionXz","SubclassXz","SubclassGPS1","SubclassGPS2"),3))
res_df=res_df%>%filter(estimator!="Index")
write.csv(res_df,file = "res_df5.csv")
```


